---
title: "Dataset testing"
author: "Lynn"
date: "3/23/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
```
```{r, include=FALSE}
#Load libraries and functions
library(dplyr)
library(data.table)
library(purrr)
library(magrittr)
source(file = "../src/parse_json_schema.R")
source(file = "../src/MIDS-calc.R")
```

## Zenodo dataset: A benchmark dataset of herbarium specimen images with label data

https://doi.org/10.5281/zenodo.3697797

Downloaded from GBIF using occ_download() based on occurrenceID or catalogNumber.

```{r}
zippath <- "../data/0176996-210914110416597.zip"
tictoc::tic()
gbif_dataset_mids <- calculate_mids(gbiffile = zippath, jsonfile = "../data/schemas/secondschema_conditions_same_level.json")

#MIDS levels
gbif_dataset_mids %>%
  group_by(MIDS_level) %>%
  tally()  %>%
  mutate(perc = n / sum(n) *100)

#MIDS achieved per condition
n_rows <- nrow(gbif_dataset_mids)
gbif_dataset_mids[ , grep("mids[0-3]", names(gbif_dataset_mids)), with = FALSE] %>% 
  map(~{(sum(.x, na.rm = TRUE) / n_rows)*100}) %>%
  as.data.table()

cat("Number of records:", n_rows)
tictoc::toc(log = TRUE, quiet = TRUE)
print(tictoc::tic.log(format = TRUE))
tictoc::tic.clearlog()
```

## BGM dataset
All specimens from the Meise Botanic Garden Herbarium (BR) on GBIF

https://doi.org/10.15468/dl.4bstmp

```{r}
zippath <- "../data/0165952-210914110416597.zip"
tictoc::tic()
gbif_dataset_mids <- calculate_mids(gbiffile = zippath, jsonfile = "../data/schemas/secondschema_conditions_same_level.json")

#MIDS levels
gbif_dataset_mids %>%
  group_by(MIDS_level) %>%
  tally()  %>%
  mutate(perc = n / sum(n) *100)

#MIDS achieved per condition
n_rows <- nrow(gbif_dataset_mids)
gbif_dataset_mids[ , grep("mids[0-3]", names(gbif_dataset_mids)), with = FALSE] %>% 
  map(~{(sum(.x, na.rm = TRUE) / n_rows)*100}) %>%
  as.data.table()

cat("Number of records:", n_rows)
tictoc::toc(log = TRUE, quiet = TRUE)
print(tictoc::tic.log(format = TRUE))
tictoc::tic.clearlog()
```

## University of Vienna, Institute for Botany - Herbarium WU

https://doi.org/10.15468/dl.pesgcz

```{r}
zippath <- "../data/0191699-210914110416597.zip"
tictoc::tic()
gbif_dataset_mids <- calculate_mids(gbiffile = zippath, jsonfile = "../data/schemas/secondschema_conditions_same_level.json")

#MIDS levels
gbif_dataset_mids %>%
  group_by(MIDS_level) %>%
  tally()  %>%
  mutate(perc = n / sum(n) *100)

#MIDS achieved per condition
n_rows <- nrow(gbif_dataset_mids)
gbif_dataset_mids[ , grep("mids[0-3]", names(gbif_dataset_mids)), with = FALSE] %>% 
  map(~{(sum(.x, na.rm = TRUE) / n_rows)*100}) %>%
  as.data.table()

cat("Number of records:", n_rows)
tictoc::toc(log = TRUE, quiet = TRUE)
print(tictoc::tic.log(format = TRUE))
tictoc::tic.clearlog()
```
## Royal Belgian Institute of Natural Sciences Bird collection + Fish collection + Mammal collection
Royal Belgian Institute of Natural Sciences Bird collection Royal Belgian Institute of Natural Sciences Fish collection Royal Belgian Institute of Natural Sciences Mammal collection

https://doi.org/10.15468/dl.uja7fq

```{r}
zippath <- "../data/0191702-210914110416597.zip"
tictoc::tic()
gbif_dataset_mids <- calculate_mids(gbiffile = zippath, jsonfile = "../data/schemas/secondschema_conditions_same_level.json")

#MIDS levels
gbif_dataset_mids %>%
  group_by(MIDS_level) %>%
  tally()  %>%
  mutate(perc = n / sum(n) *100)

#MIDS achieved per condition
n_rows <- nrow(gbif_dataset_mids)
gbif_dataset_mids[ , grep("mids[0-3]", names(gbif_dataset_mids)), with = FALSE] %>% 
  map(~{(sum(.x, na.rm = TRUE) / n_rows)*100}) %>%
  as.data.table()

cat("Number of records:", n_rows)
tictoc::toc(log = TRUE, quiet = TRUE)
print(tictoc::tic.log(format = TRUE))
tictoc::tic.clearlog()
```
## Naturalis Biodiversity Center (NL) - Paleontology

https://doi.org/10.15468/dl.n36rpw

```{r}
zippath <- "../data/0191708-210914110416597.zip"
tictoc::tic()
gbif_dataset_mids <- calculate_mids(gbiffile = zippath, jsonfile = "../data/schemas/secondschema_conditions_same_level.json")

#MIDS levels
gbif_dataset_mids %>%
  group_by(MIDS_level) %>%
  tally()  %>%
  mutate(perc = n / sum(n) *100)

#MIDS achieved per condition
n_rows <- nrow(gbif_dataset_mids)
gbif_dataset_mids[ , grep("mids[0-3]", names(gbif_dataset_mids)), with = FALSE] %>% 
  map(~{(sum(.x, na.rm = TRUE) / n_rows)*100}) %>%
  as.data.table()

cat("Number of records:", n_rows)
tictoc::toc(log = TRUE, quiet = TRUE)
print(tictoc::tic.log(format = TRUE))
tictoc::tic.clearlog()
```

##  Natural History Museum (London) Collection Specimens

https://doi.org/10.15468/dl.esgdht

```{r}
zippath <- "../data/0184293-210914110416597.zip"
tictoc::tic()
gbif_dataset_mids <- calculate_mids(gbiffile = zippath, jsonfile = "../data/schemas/secondschema_conditions_same_level.json")

#MIDS levels
gbif_dataset_mids %>%
  group_by(MIDS_level) %>%
  tally()  %>%
  mutate(perc = n / sum(n) *100)

#MIDS achieved per condition
n_rows <- nrow(gbif_dataset_mids)
gbif_dataset_mids[ , grep("mids[0-3]", names(gbif_dataset_mids)), with = FALSE] %>%
  map(~{(sum(.x, na.rm = TRUE) / n_rows)*100}) %>%
  as.data.table()

cat("Number of records:", n_rows)
tictoc::toc(log = TRUE, quiet = TRUE)
print(tictoc::tic.log(format = TRUE))
tictoc::tic.clearlog()
```

## Remarks
Certain MIDS elements included in our current MIDS interpretation seem to give recurring problems, with large numbers of specimens not meeting them:

### Modified
Despite looking in the metadata for a modified date, there are still datasets that don't meet the "modified" MIDS     element, and thus don't reach the MIDS level 0.

### MaterialType
Checks for 3 properties, which must all be true. "preparations" and "type" both seem to be absent quite often.

### TypeStatus
Current consensus is to leave this field empty when a specimen is not a type. But since in our implemenation type status is required to reach MIDS level 2, all non type specimens get stuck at MIDS level 1. Perhaps MIDS should make a recommendation on what information to put here if type status was checked, but the specimen is not a type.  

### Quality Assertations
Is almost never met. GBIF checks such a wide range of issues, that almost always something is found. Not all issues are equally important however. Perhaps we could check against a subset of issues instead of against all possible issues.
